---
title: "P8105_HW1_KJF2152.Rmd"
author: "Kaleb J. Frierson"
date: "2024-09-14"
output: 
  github_document:
    toc: true
---

# Library Calling

```{r Libraries}
library(ggplot2)


```


# Problem 1

## Loading Data

The "penguins" dataset is within the "palmerpenguins" package which I installed in the console. In the below chunk I load the dataset using the data function. 


```{r Dataset}

data("penguins", package = "palmerpenguins")

```

## Viewing Data

In the next chunk I will use the heads function to view the top few rows and all of the columns, the nrow function to get a count of rows, the ncol function to get a count of columns, and the summary function to see sumary statistics for each of the variables. 

```{r Viewing Data}

head(penguins)
nrow(penguins)
ncol(penguins)
summary(penguins)

```
## Describing Data

The penguin dataset has 344 rows each representing an individual penguin. The 8 columns include the following variables: Species, Island, Bill Length in mm, Bill Depth in mm, Flipper Length in mm, Body Mass in g, Sex, and Year of recording. 

The mean flipper length is 200.9 mm; however, the distribution of flipper length is multimodal. Thus, the Median (197.0) might be a better measure of central tendency. I determined this by graphing the variable with a histogram in the chunk below. 

```{r Describing}

ggplot(penguins, aes(x=flipper_length_mm))+ geom_histogram()

```

## Visualizing Data

In the following chunk I create a scatterplot of flipper_length_mm (y) vs bill_length_mm (x) and color points by species.

```{r Plotting}

ggplot(penguins,aes(x=bill_length_mm, y=flipper_length_mm, color=species))+
  geom_point()+
  labs(x= "Penguin Bill Length (mm)", y= "Penguin Flipper Length (mm)")

```
## Saving Plot

Using ggsave I tell R to save this plot as a png file within the figures folder of this project. 

```{r Saving}
ggsave("Figures/penguin_plot.png")

```

