---
title: "P8105_HW1_KJF2152.Rmd"
author: "Kaleb J. Frierson"
date: "2024-09-14"
output: 
  github_document:
    toc: true
---

# Library Calling

```{r Libraries}

library(ggplot2)
library(tidyverse)

```


# Problem 1

## Loading Data

The "penguins" dataset is within the "palmerpenguins" package which I installed in the console. In the below chunk I load the dataset using the data function. 


```{r Dataset}

data("penguins", package = "palmerpenguins")

```

## Viewing Data

In the next chunk I will use the heads function to view the top few rows and all of the columns, the nrow function to get a count of rows, the ncol function to get a count of columns, and the summary function to see sumary statistics for each of the variables. 

```{r Viewing Data}

head(penguins)
nrow(penguins)
ncol(penguins)
summary(penguins)

```
## Describing Data

The penguin dataset has 344 rows each representing an individual penguin. The 8 columns include the following variables: Species, Island, Bill Length in mm, Bill Depth in mm, Flipper Length in mm, Body Mass in g, Sex, and Year of recording. 

The mean flipper length is 200.9 mm; however, the distribution of flipper length is multimodal. Thus, the Median (197.0) might be a better measure of central tendency. I determined this by graphing the variable with a histogram in the chunk below. 

```{r Describing}

ggplot(penguins, aes(x=flipper_length_mm))+ geom_histogram()

```

## Visualizing Data

In the following chunk I create a scatterplot of flipper_length_mm (y) vs bill_length_mm (x) and color points by species.

```{r Plotting}

ggplot(penguins,aes(x=bill_length_mm, y=flipper_length_mm, color=species))+
  geom_point()+
  labs(x= "Penguin Bill Length (mm)", y= "Penguin Flipper Length (mm)")

```
## Saving Plot

Using ggsave I tell R to save this plot as a png file within the figures folder of this project. 

```{r Saving}
ggsave("Figures/penguin_plot.png")

```

# Problem 2

This problem is intended to emphasize variable types and introduce coercion; some awareness of how R treats numeric, character, and factor variables is necessary for working with these data types in practice.

Create a data frame comprised of:

a random sample of size 10 from a standard Normal distribution
a logical vector indicating whether elements of the sample are greater than 0
a character vector of length 10
a factor vector of length 10, with 3 different factor “levels”

## Making Data Frame

Here I create a random sample of 10 from a normal distribution (rnorm) and call it dfsample:

```{r sample}
dfsample <- rnorm(10)
```

Here I create logical, character, and factor vectors. I asked for suggestions from ChatGPT which is where I got the idea for fruit in the character vector.

```{r vectors}
logical_vector <- dfsample > 0

character_vector <- c("apple", "banana", "cherry", "date", "elderberry", 
                      "fig", "grape", "honeydew", "kiwi", "lemon")

factor_vector <- factor(c("Level1", "Level2", "Level3", "Level1", "Level2", 
                          "Level3", "Level1", "Level2", "Level3", "Level1"))
```

Here I create a data frame using the sample and vectors that I made above: 

```{r combine}
df_hw <- data.frame(RandomSample = dfsample,
                 GreaterThanZero = logical_vector,
                 CharacterVector = character_vector,
                 FactorVector = factor_vector)
df_hw
```

## Taking the Mean

Try to take the mean of each variable in your dataframe. What works and what doesn’t?

The mean of the random samples is -0.15. I obtained this by pulling the variable, calling it rs_mean, and then taking the mean of it. You could so this all in one line as mean(pull(df_hw, Random Sample)).I don't like how that looks.

```{r rs_mean}
rs_mean <- pull(df_hw, RandomSample)
mean(rs_mean)
```

Interestingly the following code works to obtain the mean of the GreaterThanZero column which is simply either true or false indicating whether the value is greater than zero. I am unsure how to interpret such a mean. 

```{r logic_mean}
logic<- pull(df_hw, GreaterThanZero)
mean(logic)
```

The following doesn't compute an average because you cannot get an average from a non-numeric vector. 

```{r character_mean, warning=FALSE}
character<- pull(df_hw, CharacterVector)
mean(character)
```

The following doesn't allow for me to obtain the mean of the factor vector. This makes sense too. 

```{r factor_mean, warning=FALSE}
factor <- pull(df_hw, FactorVector)
mean(factor)
```

## Converting Variables

In some cases, you can explicitly convert variables from one type to another. Write a code chunk that applies the as.numeric function to the logical, character, and factor variables (please show this chunk but not the output). What happens, and why? Does this help explain what happens when you try to take the mean?

```{r, echo=TRUE, message = FALSE, warning = FALSE, results='hide'}

df_hw_mutated = 
  df_hw |> 
  mutate(GreaterThanZero = as.numeric(GreaterThanZero), 
         CharacterVector = as.numeric(CharacterVector),
         FactorVector = as.numeric(FactorVector),
         )

```

You are able to turn the logical and factor vectors into numbers. The character value became NA because there was no sensible way for R to turn it into a number. When I tried to take the mean of the logical vector, I got a value that didn't make sense to me because I didn't think it was a number at all. But, R has coded it as 0 or 1 True or False and maybe in that way I was able to get an average. 

















