---
title: "P8105_HW1_KJF2152.Rmd"
author: "Kaleb J. Frierson"
date: "2024-09-14"
output: 
  github_document:
    toc: true
---

# Library Calling

```{r Libraries}

library(ggplot2)
library(tidyverse)

```


# Problem 1

## Loading Data

The "penguins" dataset is within the "palmerpenguins" package which I installed in the console. In the below chunk I load the dataset using the data function. 


```{r Dataset}

data("penguins", package = "palmerpenguins")

```

## Viewing Data

In the next chunk I will use the heads function to view the top few rows and all of the columns, the nrow function to get a count of rows, the ncol function to get a count of columns, and the summary function to see sumary statistics for each of the variables. 

```{r Viewing Data}

head(penguins)
nrow(penguins)
ncol(penguins)
summary(penguins)

```
## Describing Data

The penguin dataset has 344 rows each representing an individual penguin. The 8 columns include the following variables: Species, Island, Bill Length in mm, Bill Depth in mm, Flipper Length in mm, Body Mass in g, Sex, and Year of recording. 

The mean flipper length is 200.9 mm; however, the distribution of flipper length is multimodal. Thus, the Median (197.0) might be a better measure of central tendency. I determined this by graphing the variable with a histogram in the chunk below. 

```{r Describing}

ggplot(penguins, aes(x=flipper_length_mm))+ geom_histogram()

```

## Visualizing Data

In the following chunk I create a scatterplot of flipper_length_mm (y) vs bill_length_mm (x) and color points by species.

```{r Plotting}

ggplot(penguins,aes(x=bill_length_mm, y=flipper_length_mm, color=species))+
  geom_point()+
  labs(x= "Penguin Bill Length (mm)", y= "Penguin Flipper Length (mm)")

```
## Saving Plot

Using ggsave I tell R to save this plot as a png file within the figures folder of this project. 

```{r Saving}
ggsave("Figures/penguin_plot.png")

```

# Problem 2

This problem is intended to emphasize variable types and introduce coercion; some awareness of how R treats numeric, character, and factor variables is necessary for working with these data types in practice.

Create a data frame comprised of:

a random sample of size 10 from a standard Normal distribution
a logical vector indicating whether elements of the sample are greater than 0
a character vector of length 10
a factor vector of length 10, with 3 different factor “levels”
Try to take the mean of each variable in your dataframe. What works and what doesn’t?

Hint: to take the mean of a variable in a dataframe, you need to pull the variable out of the dataframe. Try loading the tidyverse and using the pull function.

In some cases, you can explicitly convert variables from one type to another. Write a code chunk that applies the as.numeric function to the logical, character, and factor variables (please show this chunk but not the output). What happens, and why? Does this help explain what happens when you try to take the mean?




















